### 2023-01-25 article.tex
Read all feedback
Use Springer style more
Follow Springer guidelines
Definitions below abstract
Remove table of recommendations
EOD

\begin{abstract}

@@

\end{abstract}

@@

Two different researchers write two different papers,
that are both accepted. One researcher does not supply the
computer code that was used to generate the results.
The experiment of the other researcher can be re-done easily
and you can verify that, indeed, the results are the same
as in the paper.
Do you trust the conclusion of their papers equally?

In this paper, the definition of paradata
is 'information about how data were generated',
as shown schematically in the left panel of Figure \ref{fig:figure_1}.
The term 'paradata' does not have a clear definition yet (hence this book). 
Common elements in the definitions of paradata are 
(1) paradata are usually not the primary data, (2) paradata describe
a process, (3) paradata describe the way of collecting primary data, or 
(4) paradata describe the way the primary data are generated.

@@

% \paragraph{Simple example of paradata}

A simple example would be data at eBird (\url{https://ebird.org/}),
where people can submit wher a bird of a certain species was spotted.
The name of the person that spotted that bird is paradata, 
as it is that person that generated the data and determined of
which species the bird was, 
as shown schematically in the middle panel of Figure \ref{fig:figure_1}.

% \paragraph{Relevance of paradata}

Paradata are relevant, as it allows primary data to be re-used.
In the eBird example, it helps assess the trustworthiness of the 
data, as beginner birdwatchers tend to make more mistakes in identifying
the correct bird species, as compared to experts.
By using the name and/or experience of the birder as a covariate,
we can re-use the dataset in other contexts as well,
as now this example can be generalized to
a classification problem of unlabeled
data for learning agents with different amounts 
of experience.

This paper is about paradata in the context of a computational
experiment in general and uses an example from genetic 
epidemiology in particular.
The introduction to genetic epidemiology will be postponed for now,
insteaf focussing on 'computational experiment' first.
A computational experiment is an experiment run by computer code.
In chronologic order, it usually collects primary data first, 
then runs an analysis, which finally results in the generation of results.

@@

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed \cite{vasilescu2015} and increases
the speed at which new features are added \cite{vasilescu2015}.
Using CI opens up the possibility to informally annotate features of
a repository. For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.

@@

Code may or may not be paradata, depending on how the definition
is interpreted.
Here we repeat the definition of 'paradata' and discuss 
its constituents.
This paper defined paradata as 'data about the collecting of the data'.
This implies that (1) code must be seen as data.
This paper argues, that code is data in the form of text spread
over one or more files with useful measurable properties.
(2) downloading raw data and doing calculations must be seen as collecting.
This paper argues that downloading raw datas and doing calculations, 
such as a T-test,
does describe how the bits and pieces of an end result are collected.
(3) the results of an experiment must be seen as data.
This paper argues that an experimental results is data, 
as it can be measured and can be used as the raw data of a next experiment.

@@

\rowcolors{2}{gray!10}{gray!40}
\begin{table}[h]
  \begin{tabular}{|p{2cm}|l|}
    \hline
    \textbf{Type of code} & \textbf{Recommendation} \\
    \hline
    No code       & Disallow papers that do not supply code publicly \\
    \hline
    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
    \hline
    Hosted code   & Use a standarized badge for build status \\
                  & Use a standarized badge for code coverage \\
                  & Standardize to link between paper and code host \\
    \hline
    Runnable code & Allow to upload containers without limits \\
                  & Website to host containers must allow for annotation \\
                  & Use a standarized badge for the container to be functioning  \\
                  & Standarize to link between container, paper and code repository \\
    \hline
  \end{tabular}
  \caption{Recommendations described in this paper}
  \label{tab:recommendations}
\end{table}

>>>>

\setcounter{chapter}{7}
\chapter{Research code is as vital paradata as the article itself}

\abstract{

@@

} % abstract

@@

Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that involve computation.
Both manuscripts are accepted by an equally prestigous journal after peer view. 
One researcher, however, does not supply the
computer code that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses the effect of supplying the computer 
code (from now on: 'code') 
and claims that code is paradata and that this paradata 
is just as important as the paper itself.

% \paragraph{The earliest definition of paradata}

Paradata are data about the data collection process \cite{choumert2019using}.
The concept of paradata (although not named as such yet) 
was introduced by \cite{couper1998measuring},
who developed a computer-assisted interview program
that, among others, records all key strokes,
measures the time used to answer each question 
and even the time the monitor is turned off.
The goal in that context was to assess and improve survey quality.

% \paragraph{Modern forms of paradata}

Nowadays, the lack of paradata can be seen as 'a drastic constraint'
in data use \cite{huvila2022improving}.
Given the stance that code to do a computational experiment is paradata, 
the effect of its absence is 
that a computational experiment cannot be redone easily.

This paper is about paradata in the context of a computational
experiment in general and uses an example from genetic 
epidemiology in particular.
The introduction to genetic epidemiology will be postponed for now,
instead focussing on 'computational experiment' first.
A computational experiment is an experiment run by computer code.
In chronologic order, it usually collects primary data first, 
then runs an analysis, which finally results in the generation of results.

@@

@@

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed \cite{vasilescu2015} and increases
the speed at which new features are added \cite{vasilescu2015}.
Using CI opens up the possibility to informally annotate features of
a repository. 
\iffalse
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

@@

% Reviewer 1: this section starts with a sentence that appears to undercut the whole theme of this chapter!
Code may or may not be paradata, depending on how the definition
is interpreted.
Here we repeat the definition of 'paradata' and discuss 
its constituents.
This paper defined paradata as 'data about the collecting of the data'.
This implies that (1) code must be seen as data.
This paper argues, that code is data in the form of text spread
over one or more files with useful measurable properties.
(2) downloading raw data and doing calculations must be seen as collecting.
This paper argues that downloading raw datas and doing calculations, 
such as a T-test,
does describe how the bits and pieces of an end result are collected.
(3) the results of an experiment must be seen as data.
This paper argues that an experimental results is data, 
as it can be measured and can be used as the raw data of a next experiment.

@@

\iffalse
\rowcolors{2}{gray!10}{gray!40}
\begin{table}[h]
  \begin{tabular}{|p{2cm}|l|}
    \hline
    \textbf{Type of code} & \textbf{Recommendation} \\
    \hline
    No code       & Disallow papers that do not supply code publicly \\
    \hline
    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
    \hline
    Hosted code   & Use a standarized badge for build status \\
                  & Use a standarized badge for code coverage \\
                  & Standardize to link between paper and code host \\
    \hline
    Runnable code & Allow to upload containers without limits \\
                  & Website to host containers must allow for annotation \\
                  & Use a standarized badge for the container to be functioning  \\
                  & Standarize to link between container, paper and code repository \\
    \hline
  \end{tabular}
  \caption{Recommendations described in this paper}
  \label{tab:recommendations}
\end{table}
\fi
